Incremental Forward Reasoning for White-Box Proof Search

### 2025-02-12 paper.tex
add possible todo
New title
xavier orcidID
Add abstract, modify keywords

\title{Stateful Forward Reasoning}

@@

\author{Jannis Limperg\orcidID{0000-0002-8861-5231} \and
  Xavier Généreux\orcidID{TODO}}

@@

\begin{abstract}
  TODO

  \keywords{Forward Reasoning \and Proof Automation \and Theorem Proving.}
\end{abstract}

@@

In many proof assistants, simple automation tactics based on tree search play a substantial role in practice: Isabelle's (TODO ref) and Rocq's (TODO ref) auto, ACL2's (TODO ref) waterfall, PVS's (TODO ref) grind, Lean's (TODO ref) Aesop (TODO ref), etc.
These tactics operate on \emph{goals} $Γ ⊢ A$, where the \emph{local context} $Γ$ is a list of assumptions and $A$ is the proposition to be proved.
Faced with an initial goal, the tactics try various user-specified \emph{rules}, which correspond to admissible inferences in the prover's logic.
A successful rule reduces the current goal to zero or more subgoals, which are solved recursively until either a full proof is found or no rule applies to any open goal or the search times out.

@@

>>>>

\title{Incremental Forward Reasoning for White-Box Proof Search}

@@

\author{Jannis Limperg\orcidID{0000-0002-8861-5231} \and
  Xavier Généreux\orcidID{0000-0003-4952-9557}}

@@

\begin{abstract}
  Many proof assistants provide automation tactics based on simple, tableau-style tree search, such as Isabelle's and Rocq's auto or Lean's Aesop.
  These tactics solve a given goal by recursively applying \emph{rules}, which return zero or more subgoals.
  In most cases, the rules themselves can be arbitrary tactics, so they can return any set of subgoals corresponding to an admissible inference in the prover's logic.

  In this setting, we consider \emph{forward rules}, which apply a given theorem, say $A → B → C$, to any goal containing hypotheses $A$ and $B$, adding $C$ as an additional hypothesis.
  When treated naively, such rules are tried on every goal encountered during the search, leading to repeated unifications of premises $A$ and $B$ with the hypotheses of each goal.
  Much of this work is redundant since goals often have similar hypotheses.

  We present an implementation of forward rules that takes advantage of these similarities.
  For each goal, we cache certain partial applications of forward rules in a custom data structure, the \emph{partial match state}.
  This enables efficient updates, so if goals have similar hypotheses, we can quickly update the partial match state rather than re-running all forward rules.

  Our technique is compatible with any search strategy and most logics.
  It has been implemented in the Aesop proof search tactic for Lean.

  \keywords{Forward Reasoning \and Forward Chaining \and Tactics \and Interactive Theorem Proving.}
\end{abstract}

@@

In many proof assistants, simple automation tactics based on tableau-style tree search play a substantial role in practice: Isabelle's (TODO ref) and Rocq's (TODO ref) auto, ACL2's (TODO ref) waterfall, PVS's (TODO ref) grind, Lean's (TODO ref) Aesop (TODO ref), etc.
These tactics operate on \emph{goals} $Γ ⊢ A$, where the \emph{local context} $Γ$ is a list of assumptions and $A$ is the proposition to be proved.
Faced with an initial goal, the tactics try various user-specified \emph{rules}, which correspond to admissible inferences in the prover's logic.
A successful rule reduces the current goal to zero or more subgoals, which are solved recursively until either a full proof is found or no rule applies to any open goal or the search times out.

@@

\xcom{Do we want to add potential improvements? Like an index closer to the naive algorithm.}
